---
title: "Visualizing Data"
author: "Erwing"
date: "2023-05-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing packages, this one is for readr which allows readtsv

## Had to update rlang with code, install.packages("rlang")

## install.packages("readr")

## Code to load readr

```{R}
library(readr)
```

## To read in trees dataframe

```{r}
trees <- read_tsv('data-raw/TREE_SURVEYS.txt',
               col_types = list(HEIGHT = col_double(),
                                AXIS_2 = col_double()))
```
## Received error when attempting to read_tsv says could not find function read_tsv

## Received feedback from the instructor that library() must be used for readr package

## Ran into issue that desired file does not exist in current working directory, fixed by switching directory

## Ran into issue "Warning: One or more parsing issues, call problems()` on your data frame for details, e.g.: dat <- vroom(...) problems(dat)"

## Managed to load in tree surveys data, can view now with, 'View(trees)'

##created object canopy_area with the $ function by first using the specific table trees and then multipluing the columns
```{r}
canopy_area <- (trees$AXIS_1*trees$AXIS_2)
```
## Used this code to check if the columns multiplied properly by printing and then manually calculating first few values

```{r}
canopy_area
```
## Had to add the canopy_area as a column to the trees data frame
```{r}
trees$canopy_area <- canopy_area
```

## Can view if the canopy was added with bottom code

```{r}
View(trees)
```

## Finally after about 30mins of attempting to subset I finally managed to. Turns out I tried subsetting by names, but for this format it required numbers representing the columns.
```{r}
subset(trees[c(1, 2, 3, 17)])
```

## Needed to create dataframe subset of trees

```{r}
trees_subset <- subset(trees[c(1, 2, 3, 17)])
```

## Can view the subset produced with View() function

```{r}
View(trees_subset)
```

## Plot(trees$canopy_area,trees$HEIGHT,pch=2,col=trees$TREATMENT) had issues

## Kept getting error, Error in plot.xy(xy, type, ...) : invalid color name 'TOTAL', to fix issue I found that I neeeded to change the TREATMENT from a character to a factor type.

## Issue was fixed with code: trees$TREATMENT = as.factor(trees$TREATMENT)

```{r}
trees$TREATMENT = as.factor(trees$TREATMENT)
```

## Had to set color with code:
```{r}
col=rgb(100,0,0,50,maxColorValue=255)
```

## This code functions as plot(x axis, y axis, pch(which is point size), col(defined color) = (set variable for distinguishing data))

```{r}
plot(trees$canopy_area,trees$HEIGHT,pch=2,col=trees$TREATMENT, xlab = "Canopy Area (m)" , ylab = "Height (m)", main = "Acacia Height vs Canopy") 
```

## Read in ggplot2

```{r}
library(ggplot2)
```

## Now use the code below to make subplots

```{R}
ggplot(data = trees, mapping = aes(x = canopy_area, y = HEIGHT, color = TREATMENT)) + geom_point() +
labs (x = "Canopy area") +
facet_wrap(~SPECIES)
```


## Exercise 4

## Load in txt file
```{r}
acacia <- read.csv(file = "/Users/sanch/Documents/Schooldownloads/BIO197/Rstudio/Homeworks/Homeworks/data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt",
                   sep ="\t",
                   na.strings = "dead")
is.numeric(acacia$HEIGHT)
```
```{r}
ggplot() +
geom_point(data = trees, mapping = aes(x= CIRC, y= HEIGHT), color = "gray", alpha = 0.5) +
geom_point(data = acacia, mapping =aes (x = AXIS1, y=AXIS2), color = "red", alpha = 0.6)
```








